---
title: "LanguageAIProject"
output: html_document
date: "2025-03-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

##Loading packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lme4)
```


##The prepping of the data



```{r}
#The Output_Full_withNUll file has null data for the demographic categories. This was to allow a comparison between the null persona and other personas using a mixed-effects model
aidfwithnull <- read.csv("Output_Full_withNull.csv")
aidfwithnull <- mutate(aidfwithnull, relent = outentail/120*100)
aidfwithnull <- mutate(aidfwithnull, relcon = outcon/120*100)
aidfwithnull <- mutate(aidfwithnull, relneu = outneu/120*100)
aidfwithnull <- mutate(aidfwithnull, dif = (abs(inneu-relneu)+abs(inent-relent)+abs(incon-relcon))/2)
aidfwithnull$age <- as.character(aidfwithnull$age)

#Output_Full is a csv file with blank slots for the null persona's demographic information
aidf <- read.csv("Output_Full.csv")
aidf <- mutate(aidf, relent = outentail/120*100)
aidf <- mutate(aidf, relcon = outcon/120*100)
aidf <- mutate(aidf, relneu = outneu/120*100)
aidf <- mutate(aidf, dif = (abs(inneu-relneu)+abs(inent-relent)+abs(incon-relcon))/2)

```



##Checking the null persona
The following code checks models where null values for the demographic categories were added in the final output file. This is to make sure that we can compare the dif. scores (absolute difference between NLI predictions of humans and LLM) with the null persona.




```{r}

#These models below were to check whether the persona can be a predictor for the dif score.
modelpeople = lmer(dif ~1+ (1|persona_nr) + (1|inentro), data = aidf)
summary(modelpeople)
modelwithoutpers = lmer(dif ~1 + (1|inentro), data = aidf)
anova(modelpeople, modelwithoutpers)
modelpeoplefixef = lmer(dif ~1+ persona_nr + (1|inentro), data = aidf)
summary(modelpeoplefixef)

anova(modelpeoplefixef, modelwithoutpers)


#These models below were again to check but then with null data in the demographic slots.



modelallwithnull = lmer(dif~1+ gender + nationality + education + age + (1+ gender + nationality + education + age|inentro), data = aidfwithnull)
summary(modelallwithnull)
modelonlyrandwithnull =lmer(dif~1+ (1+ gender + nationality + education + age|inentro), data = aidfwithnull)
anova(modelonlyrandwithnull, modelallwithnull)
summary(modelonlyrandwithnull)
modelwoagenatwithnull = lmer(dif~1+ education + gender + (1+ gender + nationality + education + age|inentro), data = aidfwithnull)
summary(modelwoagenatwithnull)
anova(modelwoagenatwithnull, modelonlyrandwithnull)

modelwoagenatinterwithnull = lmer(dif~1+ education + gender + education*gender + (1+ gender + nationality + education + age|inentro), data = aidfwithnull)
summary(modelwoagenatinterwithnull)
anova(modelwoagenatinterwithnull, modelwoagenatwithnull)


```


##Checking differences between entropy levels


```{r}

#To check for the influence of entropy on the dif score

mentro = lm(dif~1+ entropylevel, data = aidf)
summary(mentro)
mentro2 = lm(dif ~1+ inentro, data= aidf)
summary(mentro2)
anova(mentro, mentro2)



```







##Checking differences between persona without null persona



```{r}


#To visualise the data
hist(aidf$dif)
aidf %>% ggplot(aes(x=dif, y=inentro))+geom_point(col= "plum")
boxplot(dif ~inentro, col = "plum", aidf)
boxplot(dif ~gender, col = "plum", aidf)
boxplot(dif ~nationality, col = "plum", aidf)
boxplot(dif ~education, col = "plum", aidf)
boxplot(dif ~age, col = "plum", aidf)
boxplot(dif ~gender*nationality, col="plum", aidf)

#To remove the data for null persona
aidf2 <- aidf %>% filter(persona_nr != 1)


#This is model comparison
modelall = lmer(dif~1+ gender + nationality + education + age + (1+ gender + nationality + education + age|inentro), data = aidf2)
summary(modelall)
modelonlyrand =lmer(dif~1+ (1+ gender + nationality + education + age|inentro), data = aidf2)
anova(modelonlyrand, modelall)

m1 = lmer(dif~1+ education + gender + (1+ gender + nationality + education + age|inentro), data = aidf2)
summary(m1)
anova(m1, modelonlyrand)

mfinal = lmer(dif~1+ education + gender + education*gender + (1+ gender + nationality + education + age|inentro), data = aidf2)
summary(m4)
anova(mfinal, m1)

m4 = lmer(dif~1+ education + gender + education*gender + (1+ gender + education |inentro), data = aidf2)
anova(m1, m4)
anova(mfinal, m4)

#These anovas above show that 'mfinal' is the best fitting model

#Visualisation
boxplot(dif ~gender*education, col="plum", aidf)

summary(mfinal)



# Ensure correct factor order
aidf2$entrolevel <- factor(aidf2$entrolevel, levels = c("Low", "Medium", "High"))
aidf2$education <- factor(aidf2$education, levels = c("highschool", "college", "graduateschool"))

# Create the boxplot
ggplot(aidf2, aes(x = gender, y = dif, fill = education)) +
  geom_boxplot(position = position_dodge(width = 0.8)) +  # Dodge to separate gender groups
  facet_grid(. ~ entrolevel) +  # Gender facets separately
  labs(title = "Difference Score by Gender and Education",
       x = "Entropy Level", y = "DIF Score") +
  scale_fill_manual(values = c("highschool" = "pink", "college" = "lightblue", "graduateschool" = "darkseagreen2")) +
  theme_minimal()



```

## Model comparison steps

Below are the different models we compared with each other. 

```{r}
model1 = lmer(dif~ 1+ (1|inentro), data = aidf)
print(summary(model1))
model2.1 = lmer(dif~ 1+ gender + (1|inentro), data = aidf)
print(summary(model2))
model3 = lmer(dif~ 1+ gender + nationality + (1|inentro), data = aidf)
print(summary(model3))
model4 = lmer(dif~ (1|inentro)+gender+nationality+education, data = aidf)
print(summary(model4))
model5= lmer(dif~ (1|inentro)+gender+nationality+education+age, data=aidf)
summary(model5)

modela = lmer(dif~1+ gender + nationality + education + age + (1+ gender + education + age|inentro), data = aidf)

anova(modela, modelall)

modelwofe = lmer(dif~1+ gender + nationality + education + age + (1|inentro), data = aidf)
anova(modelwofe, modelall)



modelwoage = lmer(dif~1+ gender + nationality + education + (1+ gender + nationality + education + age|inentro), data = aidf)
anova(modelwoage, modelonlyrand)

modelwoge = lmer(dif~1 + nationality + education + age + (1+ gender + nationality + education + age|inentro), data = aidf)
anova(modelwoge, modelonlyrand)

modelwonat = lmer(dif~1 + education + age + gender + (1+ gender + nationality + education + age|inentro), data = aidf)
anova(modelwonat, modelonlyrand)

modelwoned = lmer(dif~1 + nationality + age + gender + (1+ gender + nationality + education + age|inentro), data = aidf)
anova(modelwoned, modelonlyrand)



```












